Distributional Trees and Forests

### 2017-11-08 pkg/disttree/inst/draft/draft.Rnw
edited plots and text in section 'simulation' and restructured section 'probabilistic forcasting'
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1686 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: bda48e60a0ced43e0b039206cc583a72ab3beca8

\section{Introduction}
A typical application in probabilistic forecasting in meteorology is the analysis of precipitation data. The two main interests lie in estimating the probability of rain as well as predicting the expected amount of precipitation. In this research area various methods have been developed and tested in the last decades. (Multiple) Linear models such as ensemble model output statistics (EMOS) were introduced more than a decade ago (\cite{gneiting2005calibrated}) and are still common and widely used. Despite their simplicity and interpretability, the limitation to modeling only linear effects has risen the interest in applying more flexible models. This feature is provided by generalized additive models (GAM) which can also model non-linear effects. However, often only the expectation of the underlying distribution is estimated. To obtain probabilistic predictions, generalized additive models for location, scale, and shape (GAMLSS, \cite{stasinopoulos2005}) offer a framework in which each distribution parameter can be modeled separately by its own GAM. In Figure \ref{Devel_parmod} the development of parametric models starting from simple linear models (LM) and ranging to GAMLSS is illustrated.\\
These flexible parametric methods provide well fitting models when applied on precipitation data in a rather smooth/homogeneous setting, e.g. areas where no abrupt changes in meteorological/climatological events appear. However, for example in complex terrain/mountainous regions various factors can change rapidly and their effects are difficult to be modeled/represented by a smooth parametric model. In this situation a different approach which can deal with these abrupt changes would be to apply tree structured models such as regression trees (\cite{loh2011classification}) or random forests (\cite{breiman2001random}). Moreover, these models are particularly advantageous in case of many covariates with unknown effects or interactions as the variable selection is done automatically within the model fitting process. Therefore, no additional effort or knowledge for the specification of the model in advance is required. However, even though their main advantage of being able to model non-additive effects and interactions without any necessary prespecification can play to their strength, regression trees are not an adequate choice if (additional) smooth effects appear as well. In this situation a random forest provides a good compromise as it combines a set of trees such that the steps imposed by the trees are smoothed out in case of smooth effects while abrupt changes are still captured by the tree structure of the model.\\
For these reasons, regression trees and particularly random forests provide valuable features which can contribute to improving models for probabilistic forecasting. But, up to now, no general distributional framework is available for these methods. Therefore, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. The developments in the field of tree-structured models together with the novel approaches of distributional trees and forests are illustrated in Figure \ref{Devel_treeforest}.\\
After explaining the novel methodology in section \ref{Methodology} of this paper and presenting results from a simulation study in section \ref{Simulation} the focus will then be put on an application in the field of probabilistic forecasting in section \ref{ProbabilisticForecasting}. The new method is applied on precipitation data in a mountainous region with a large number of numerical weather prediction quantities and the results are compared to those of other methods commonly used in this research area. In a first setting only one observation station is taken into consideration and based on the data/information from 25 years probabilistic predictions are made for the 3 successive years. A brief outlook on the results of this setting is given by Figure \ref{axams_onestation}. For the 24th of July of 2010, 2011 and 2012 the density functions of the estimated distributions for this day of all three prediction years are plotted together with the corresponding observations. The estimated distribution parameters for the 24th of July of 2010 are comparably high with location $3.146$ and scale $1.902$ which leads to a well fitting predicted expectation of $3.185$ close to the observed value of $3.001$. Also for 2011 and 2012 the estimated expected values of 1.085 and 0.483 provide reasonable predictions as the observed values are 1.121 and 0 respectively.\\
Next to the expected values the predictions provided by these distributional models offer a wide range of information as they specify the whole distribution and therefore enable further methods of inference and analysis.

@@

\subsection{Motivation (Review/Problems)}
As illustrated in Figure \ref{Devel_parmod} a lot of developement has been made in the field of parametric modeling and a wide range of tools can be applied in research areas such as probabilistic forecasting. But as mentioned before, further features such as the ability to capture non-additive effects or interactions are desirable in many applications which demands for more flexible models. Since tree structured models provide these required features a combination of them with parametric models can offer a framework that is able to deal with a wide range of different modeling situations. In this section we are going to introduce the novel methodology which embeds parametric modeling into the idea of tree structered models step by step starting with a simple distributional fit in subsection \ref{distfit}, then building a distributional tree in subsection \ref{disttree} and finally a distributional forest in subsection \ref{distforest}.

@@

The goal is to fit a distributional model $D(Y, \theta)$ to a response variable $Y \in \mathcal{Y}$ for a \textit{k}-dimensional vector of distribution parameters $\theta \in \Theta$ in case of a \textit{k}-parametric distribution. The distribution family has to be specified in advance such that a log-likelihood function $\ell(\theta; Y)$ is provided. Using the GAMLSS framework (\cite{stasinopoulos2005}) a wide range of distributions can be selected where each distribution parameter can be specified separatly. This enables the model to deal with various features that can usually be very challenging in model fitting such as censoring, truncation or heavy tails.\\

@@

The following methods are applied and compared to distributional trees and forests:
\begin{itemize}
\item gamlss \cite{stasinopoulos2005} fits a separate generalized additive model to each of the distribution models. In that way it specifies a whole distribution and is a very efficient option for smooth effects of the covariates on the parameters.
\item bamlss \cite{bamlss2017} also provides an infrastructure for estimating probabilistic distributional regression but within a Bayesian Framework. Again each parameter is modeled by (possibly) complex additive terms.
\item gamboostLSS \cite{gamboostLSS2017} offers boosting models for fitting  generalized additive models for location, shape and scale
\item randomForest \cite{randomForest2002} combines an ensemble of classification and regression trees applying bagging and reducing the correlation among trees through a randomized selection of possible split variables within each node.
\item cforest \cite{hothorn2015partykit} builds a forest in a similar way but consisting of ctrees applying the ctree algorithm mentioned in section \ref{Methodology}. In this simulation study no transformation function is included.
\end{itemize}

The different methods are tested on generated data sets consisting of the values of a response variable $Y$ and ten covariates $X_1, \ldots, X_{10}$ which are all independent und identically distributed following a uniform distribution.
$$
X_i \sim \textit{Unif}([-1,1]) \qquad \forall i \in \{1,\ldots,10\}
$$

The distribution of the response variable is a left censored normal distribution with censoring point 0. The corresponding distribution parameters are defined by a parameter function $f$ depending on the covariates and an additional factor $\kappa$ which regulates the steepness of certain sections of the parameter function. For low values of $\kappa$ the resulting parameters change rather slowly over varying values of the covariates while high values of $\kappa$ lead to abrupt changes. \\ 
The location parameter $\mu$ depends on $X_1$ and $\kappa$ and the scale parameter $\sigma$ depends on $X_2$ and $\kappa$. The other covariates $X_3, \ldots, X_{10}$ are used as noise variables.\\
The latent variable $Y^{\ast}$ is generated based on the following setting.
\begin{center}
$$ 
Y^{\ast} \sim \mathcal{N} (f(X_1, X_2, \kappa))
$$
with
$$
f(X_1, X_2, \kappa) = \left(\mu(X_1,\kappa) , \sigma(X_2,\kappa)\right)
$$ 
and
$$
\mu(X_1,\kappa) = \mu_{\text{base}} + 8 \cdot \exp{(-(3 \cdot X_1-1)^{(2\cdot \kappa)})}
$$

$$
\sigma(X_2,\kappa) = 2 + 2 \cdot (1-\Lambda(\kappa^{1.8} \cdot 5 \cdot (X_2+0.5)))
$$
where $\Lambda$ is the cumulative distribution function of the logistic distribution
$$
\Lambda(\omega) = \frac{1}{1 + \exp(-\omega)}
$$
The response variable $Y$ is defined as 
$$
Y = 
\begin{cases}
    Y^{\ast},& \text{if } Y^{\ast} \geq 0\\
    0,       & \text{if } else
\end{cases}
$$
\end{center}

To illustrate the parameter functions and the impact of $\kappa$ together with the results of the considered methods they are first applied to one generated data set only where $\kappa$ is set to 1 (top plots) and then to a second data set where $\kappa$ is set to 10 (bottom plots).\\
In the left plots the location parameter function $\mu(X_1, \kappa)$ is plotted along the covariate $X_1$ together with the estimated values of the different models for fixed covariates $(x_2, x_3, \ldots, x_{10}) = (0,0, \ldots, 0)$. \\
In the right plots the scale parameter function $\sigma(X_2, \kappa)$ is plotted along the covariate $X_2$ together with the estimated values of the different models for fixed covariates $(x_1, x_3, \ldots, x_{10}) = (0.4, 0, \ldots, 0)$. \\

@@

<<echo=FALSE, results=hide>>=
source("oneset.R")
if(file.exists("oneset1.rda")){
  load("oneset1.rda")
} else {
  oneset1 <- sim_oneset(kappa = 7, nobs = 400,
                       seedconst = 7, ntree = 100,
                       formula = y~x1+x2+x3+x4+x5+x6, 
                       tree_minsplit = 25, tree_minbucket = 10, tree_mincrit = 0.95, 
                       forest_minsplit = 25, forest_minbucket = 10, forest_mincrit = 0, 
                       forest_mtry = 3,
                       type.tree = "ctree",
                       censNO = TRUE,
                       fix.mu = FALSE,
                       fix.sigma = FALSE,
                       mu.sigma.interaction = FALSE,
                       pred_fix = TRUE,
                       pred_fix_x1 = -0.5,
                       pred_fix_x4 = 0,
                       gamboost_cvr = FALSE,
                       eval_disttree = TRUE,
                       eval_distforest = TRUE,
                       eval_gamlss = TRUE,
                       eval_gamboostLSS = FALSE,
                       eval_randomForest = FALSE,
                       eval_cforest = FALSE,
                       mubase = 3,
                       sigmabase = 3)
  save(oneset1, file = "oneset1.rda")
}

if(file.exists("oneset10.rda")){
  load("oneset10.rda")
} else {
  oneset10 <- sim_oneset(kappa = 10, nobs = 400,
                       seedconst = 7, ntree = 100,
                       formula = y~x1+x2+x3+x4+x5+x6, 
                       tree_minsplit = 25, tree_minbucket = 10, tree_mincrit = 0.95, 
                       forest_minsplit = 25, forest_minbucket = 10, forest_mincrit = 0, 
                       forest_mtry = 3,
                       type.tree = "ctree",
                       censNO = TRUE,
                       fix.mu = FALSE,
                       fix.sigma = FALSE,
                       mu.sigma.interaction = FALSE,
                       pred_fix = TRUE,
                       pred_fix_x1 = -0.5,
                       pred_fix_x4 = 0,
                       gamboost_cvr = FALSE,
                       eval_disttree = TRUE,
                       eval_distforest = TRUE,
                       eval_gamlss = TRUE,
                       eval_gamboostLSS = FALSE,
                       eval_randomForest = FALSE,
                       eval_cforest = FALSE,
                       mubase = 3,
                       sigmabase = 3)
  save(oneset10, file = "oneset10.rda")
}
@

\begin{figure}[t!]
\begin{subfigure}{0.6\textwidth}
<<sim_oneset1_mu, fig=TRUE, echo=FALSE>>=
plot_oneset(oneset1, 
            compare_mu = TRUE, 
            add_dt = TRUE, add_df = TRUE, add_g = TRUE #, 
            #add_gb = TRUE, add_rf = TRUE, add_cf = TRUE
            )
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.6\textwidth}
<<sim_oneset1_sigma, fig=TRUE, echo=FALSE>>=
plot_oneset(oneset1, 
            compare_sigma_line = TRUE, 
            add_dt = TRUE, add_df = TRUE, add_g = TRUE #, 
            #add_gb = TRUE, add_rf = TRUE, add_cf = TRUE
            )
@
\end{subfigure}
\caption{\label{plot_sim_oneset1} ... }
\end{figure}

\begin{figure}[t!]
\begin{subfigure}{0.6\textwidth}
<<sim_oneset10_mu, fig=TRUE, echo=FALSE>>=
plot_oneset(oneset10,
            compare_mu = TRUE,
            add_dt = TRUE, add_df = TRUE, add_g = TRUE #, 
            #add_gb = TRUE, add_rf = TRUE, add_cf = TRUE
            )
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.6\textwidth}
<<sim_oneset10_sigma, fig=TRUE, echo=FALSE>>=
plot_oneset(oneset10, 
            compare_sigma_line = TRUE, 
            add_dt = TRUE, add_df = TRUE, add_g = TRUE #, 
            #add_gb = TRUE, add_rf = TRUE, add_cf = TRUE
            )
@
\end{subfigure}
\caption{\label{plot_sim_oneset10} ... }
\end{figure}

Now, to get a more profound impression the methods are evaluated on various data sets. The value of $\kappa$ is increasing from 1 to 10 by a stepsize of 1 and for each step ... data sets are generated. The methods are then compared by the out-of-sample RMSE for the location parameter $\mu$ and the shape parameter $\sigma$ as well as by the out-of-sample log-likelihood of the fitted models, averaged within each step.
(In that way it can be investigated how the results change for a changing value of $\kappa$.)

@@

\begin{figure}[t!]
\begin{subfigure}{0.6\textwidth}
<<sim_rmse_exp, fig=TRUE, echo=FALSE>>=
plot_rmse(simres, type = "exp")
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.6\textwidth}
<<sim_rmse_par, fig=TRUE, echo=FALSE>>=
plot_rmse(simres, type = "par")
@
\end{subfigure}
\caption{\label{plot_sim_rmse}Results of the simulation study (RMSE).\\
left: continuous lines represent the difference between the true and the estimated expected value, dotted lines represent the difference between the observations and the estimated expected value\\
right: continuous lines represent the location parameter, dotted lines the scale parameter}
\end{figure}

\begin{figure}[t!]
\begin{subfigure}{0.6\textwidth}
<<sim_ll, fig=TRUE, echo=FALSE>>=
plot_ll(simres)
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.6\textwidth}
<<sim_levelplot, fig=TRUE, echo=FALSE>>=
levelfun <- function(x) simres$fun(x,5)

ncuts <- 100
colR <- colorRampPalette(c("blue","green"))                                       
cols <- colR(ncuts+1)
griddata <- expand.grid(list(x1 = seq(-1, 1, 0.01), x4 = seq(0, 1, 0.2)))
griddata$mu <- levelfun(griddata)
levelplot(mu ~ x1 * x4, data = griddata, xlab = "x1", ylab = "x4", 
          region = TRUE, cuts = ncuts, col.regions = cols)

@
\end{subfigure}
\caption{\label{plot_sim_ll}left: results of the simulation study: log-likelihood\\ %\code{disttree} vs. \code{distforest} vs. \code{gamlss} vs. \code{randomForest} vs. %\code{bamlss} vs. \code{gamboostLSS}\\
right: levelplot of the parameter function for $\kappa=5$}
\end{figure}

>>>>

\SweaveOpts{eval = TRUE}
\section{Introduction}
A typical application in probabilistic forecasting in meteorology is the analysis of precipitation data. The two main interests lie in estimating the probability of rain as well as predicting the expected amount of precipitation. In this research area various methods have been developed and tested in the last decades. (Multiple) Linear models such as ensemble model output statistics (EMOS) were introduced more than a decade ago (\cite{gneiting2005calibrated}) and are still common and widely used. Despite their simplicity and interpretability, the limitation to modeling only linear effects has risen the interest in applying more flexible models. This feature is provided by generalized additive models (GAM) which can also model non-linear effects. However, often only the expectation of the underlying distribution is estimated. To obtain probabilistic predictions, generalized additive models for location, scale, and shape (GAMLSS, \cite{stasinopoulos2005}) offer a framework in which each distribution parameter can be modeled separately by its own GAM. In Figure \ref{Devel_parmod} the development of parametric models starting from simple linear models (LM) and ranging to GAMLSS is illustrated.\\
These flexible parametric methods provide well fitting models when applied on precipitation data in a rather smooth/homogeneous setting, e.g. areas where no abrupt changes in meteorological/climatological events appear. However, for example in complex terrain/mountainous regions various factors can change rapidly and their effects are difficult to be modeled/represented by a smooth parametric model. In this situation a different approach which can deal with these abrupt changes would be to apply tree structured models such as regression trees (\cite{loh2011classification}) or random forests (\cite{breiman2001random}). Moreover, these models are particularly advantageous in case of many covariates with unknown effects or interactions as the variable selection is done automatically within the model fitting process. Therefore, no additional effort or knowledge for the specification of the model in advance is required. However, even though their main advantage of being able to model non-additive effects and interactions without any necessary prespecification can play to their strength, regression trees are not an adequate choice if (additional) smooth effects appear as well. In this situation a random forest provides a good compromise as it combines a set of trees such that the steps imposed by the trees are smoothed out in case of smooth effects while abrupt changes are still captured by the tree structure of the model.\\
For these reasons, regression trees and particularly random forests provide valuable features which can contribute to improving models for probabilistic forecasting. But, up to now, no general distributional framework is available for these methods. Therefore, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. The developments in the field of tree-structured models together with the novel approaches of distributional trees and forests are illustrated in Figure \ref{Devel_treeforest}.\\

After explaining the novel methodology in Section \ref{Methodology} of this paper and presenting results from a simulation study in Section \ref{Simulation} the focus will then be put on an application in the field of probabilistic forecasting in Section \ref{ProbabilisticForecasting}. The new method is applied on precipitation data in a mountainous region with a large number of numerical weather prediction quantities and the results are compared to those of other methods commonly used in this research area. In a first setting only one observation station is taken into consideration and based on the data/information from 25 years probabilistic predictions are made for the 3 successive years. A brief outlook on the results of this setting is given by Figure \ref{axams_onestation}. For the 24th of July of 2010, 2011 and 2012 the density functions of the estimated distributions for this day of all three prediction years are plotted together with the corresponding observations. The estimated distribution parameters for the 24th of July of 2010 are comparably high with location $3.146$ and scale $1.902$ which leads to a well fitting predicted expectation of $3.185$ close to the observed value of $3.001$. Also for 2011 and 2012 the estimated expected values of 1.085 and 0.483 provide reasonable predictions as the observed values are 1.121 and 0 respectively.\\
Next to the expected values the predictions provided by these distributional models offer a wide range of information as they specify the whole distribution and therefore enable further methods of inference and analysis.

@@

\subsection{Motivation (Review/Problems)}
As illustrated in Figure \ref{Devel_parmod} a lot of development has been made in the field of parametric modeling and a wide range of tools can be applied in research areas such as probabilistic forecasting. But as mentioned before, further features such as the ability to capture non-additive effects or interactions are desirable in many applications which demands for more flexible models. Since tree structured models provide these required features a combination of them with parametric models can offer a framework that is able to deal with a wide range of different modeling situations. In this section we are going to introduce the novel methodology which embeds parametric modeling into the idea of tree structured models step by step starting with a simple distributional fit in Subsection \ref{distfit}, then building a distributional tree in Subsection \ref{disttree} and finally a distributional forest in Subsection \ref{distforest}.

@@

The goal is to fit a distributional model $D(Y, \theta)$ to a response variable $Y \in \mathcal{Y}$ for a \textit{k}-dimensional vector of distribution parameters $\theta \in \Theta$ in case of a \textit{k}-parametric distribution. The distribution family has to be specified in advance such that a log-likelihood function $\ell(\theta; Y)$ is provided. Using the GAMLSS framework (\cite{stasinopoulos2005}) a wide range of distributions can be selected where each distribution parameter can be specified separately. This enables the model to deal with various features that can usually be very challenging in model fitting such as censoring, truncation or heavy tails.\\

@@

To prove these statements distributional trees and forests are evaluated on generated data sets. Their results are then compared to those of a GAMLSS (\cite{stasinopoulos2005}) which fits a separate generalized additive model to each of the distribution models and in that way specifies a whole distribution and is a very efficient option for smooth effects of the covariates on the parameters.\\
The generated data sets consist of the values of a response variable $Y$ and six covariates $X_1, \ldots, X_6$ which are all independent and distributed as follows:
$$
X_1, X_2, X_3 \sim \textit{Unif}([-1,1])
$$
$$
X_4, X_5, X_6 \sim \textit{Binom}(1, 0.5)
$$

The distribution of the response variable is a left censored normal distribution with censoring point 0. The corresponding distribution parameters are defined by a parameter function $f$ depending on the covariates and an additional factor $\kappa$ which regulates the steepness of certain sections of the parameter function. For low values of $\kappa$ the resulting parameters change rather slowly over varying values of the covariates while high values of $\kappa$ lead to abrupt changes. \\ 
The location parameter $\mu$ depends on $X_1, X_4$ and $\kappa$ and the scale parameter $\sigma$ depends on $X_1, X_4$ and $\kappa$. The other covariates $X_2, X_3, X_5$ and $X_6$ are used as noise variables.\\
The latent variable $Y^{\ast}$ is generated based on the following setting.
\begin{center}
$$ 
Y^{\ast} \sim \mathcal{N} (f(X_1, X_4, \kappa))
$$
with
$$
f(X_1, X_2, \kappa) = \left(\mu(X_1,\kappa) , \sigma(X_2,\kappa)\right)
$$ 
and
$$
\mu(X_1, X_4, \kappa) = 0.2 + 7 \cdot \exp{(-(3 \cdot X_1-1.2)^{(2\cdot \kappa)})} \cdot \mathbf{1}(X_1 < 0.4) + 
                              7 \cdot \exp{(-(3 \cdot X_1-1.2)^{(2\cdot \kappa)})} \cdot \frac{1}{2} + \frac{7}{2} \cdot \mathbf{1}(X_1 \geq 0.4) + 
                              3 \cdot \exp{(-(3 \cdot X_1-1.2)^{(2\cdot \kappa)})} \cdot X_4
$$

$$
\sigma(X_2,\kappa) = 0.5 + \frac{\mu}{5} + 2 \cdot \mathbf{1}(X_1 < -0.7)
$$
%where $\Lambda$ is the cumulative distribution function of the logistic distribution
%$$
%\Lambda(\omega) = \frac{1}{1 + \exp(-\omega)}
%$$
The response variable $Y$ is defined as 
$$
Y = 
\begin{cases}
    Y^{\ast},& \text{if } Y^{\ast} \geq 0\\
    0,       & else
\end{cases}
$$
\end{center}

To illustrate the parameter functions and the impact of $\kappa$ together with the results of the considered methods they are first applied on one generated data set only where $\kappa$ is set to 1 (top plots in Figure \ref{}) and then to a second data set where $\kappa$ is set to 10 (bottom plots in Figure \ref{}).\\
In the left plots the location parameter function $\mu(X_1, \kappa)$ is plotted along the covariate $X_1$ together with the estimated values of the different models for fixed covariates $(x_2, x_3, \ldots, x_{10}) = (0,0, \ldots, 0)$. \\
In the right plots the scale parameter function $\sigma(X_2, \kappa)$ is plotted along the covariate $X_2$ together with the estimated values of the different models for fixed covariates $(x_1, x_3, \ldots, x_{10}) = (0.4, 0, \ldots, 0)$. \\

@@

<<echo=FALSE, results=hide>>=
source("oneset.R")
if(file.exists("oneset1.rda")){
  load("oneset1.rda")
} else {

  library("gamlss.cens")
  gen.cens("NO", type = "left")

  kappa <- 1
  fun <- function(x){
    mu <- 0.2 + 
      (10 * (exp(-(3*x[,1]-1.2)^(2*kappa)))) * as.numeric(x[,1] < 0.4) + 
      (10 * (exp(-(3*x[,1]-1.2)^(2*kappa))) * (1/2) + 10/2) * as.numeric(x[,1]>= 0.4) #+
    #(3 * (exp(-(3*x[,1]-1.2)^(2*kappa)))) +
    #(3 * (exp(-(3*x[,1]-1.2)^(2*kappa)))) * x[,4]
    sigma <- 0.5 + 
      #mu/5 + 
      2 * (x[,1] < -0.4) + 
      3 * x[,4] #+
    #1*(x[,1] < -0.7) 
    #plogis((kappa^(1.8)) * 4 * (x[,2] + 0.7)) #* x[,4]
    #4*(1-plogis((kappa^(1.8)) * 2 * (x[,1] + 0.5))) * x[,4]
    #5 * abs(x[,1]) * x[,4]
    par <- cbind(mu, sigma)
    return(par)
  }

  oneset1 <- sim_oneset(parfun = fun, nobs = 500, seedconst = 74, ntree = 100,
                        formula = y~x1+x2+x3+x4+x5+x6, 
                        #formula = y~x1+x2+x4+x5,
                        #formula = y~x1+x2+x3,
                        tree_minsplit = 20, tree_minbucket = 7, tree_mincrit = 0.95, 
                        forest_minsplit = 20, forest_minbucket = 7, forest_mincrit = 0, 
                        forest_mtry = 3, type.tree = "ctree", 
                        censNO = TRUE, 
                        pred_fix = TRUE,
                        pred_fix_x1 = 0.9,
                        pred_fix_x2 = 0.9,
                        pred_fix_x4 = 1,
                        gamboost_cvr = FALSE,
                        eval_disttree = TRUE,
                        eval_distforest = TRUE,
                        eval_gamlss = TRUE,
                        nrunif = 3,
                        nrcovform = 6)

  save(oneset1, file = "oneset1.rda")
}

if(file.exists("oneset10.rda")){Hans-Brenner-Platz
  load("oneset10.rda")
} else {

  library("gamlss.cens")
  gen.cens("NO", type = "left")

  kappa <- 10
  fun <- function(x){
    mu <- 0.2 + 
      (10 * (exp(-(3*x[,1]-1.2)^(2*kappa)))) * as.numeric(x[,1] < 0.4) + 
      (10 * (exp(-(3*x[,1]-1.2)^(2*kappa))) * (1/2) + 10/2) * as.numeric(x[,1]>= 0.4) #+
    #(3 * (exp(-(3*x[,1]-1.2)^(2*kappa)))) +
    #(3 * (exp(-(3*x[,1]-1.2)^(2*kappa)))) * x[,4]
    sigma <- 0.5 + 
      #mu/5 + 
      2 * (x[,1] < -0.4) + 
      3 * x[,4] #+
    #1*(x[,1] < -0.7) 
    #plogis((kappa^(1.8)) * 4 * (x[,2] + 0.7)) #* x[,4]
    #4*(1-plogis((kappa^(1.8)) * 2 * (x[,1] + 0.5))) * x[,4]
    #5 * abs(x[,1]) * x[,4]
    par <- cbind(mu, sigma)
    return(par)
  }

  oneset10 <- sim_oneset(parfun = fun, nobs = 500, seedconst = 74, ntree = 100,
                        formula = y~x1+x2+x3+x4+x5+x6, 
                        #formula = y~x1+x2+x4+x5,
                        #formula = y~x1+x2+x3,
                        tree_minsplit = 20, tree_minbucket = 7, tree_mincrit = 0.95, 
                        forest_minsplit = 20, forest_minbucket = 7, forest_mincrit = 0, 
                        forest_mtry = 3, type.tree = "ctree", 
                        censNO = TRUE, 
                        pred_fix = TRUE,
                        pred_fix_x1 = 0.9,
                        pred_fix_x2 = 0.9,
                        pred_fix_x4 = 1,
                        gamboost_cvr = FALSE,
                        eval_disttree = TRUE,
                        eval_distforest = TRUE,
                        eval_gamlss = TRUE,
                        nrunif = 3,
                        nrcovform = 6)

  save(oneset10, file = "oneset10.rda")
}
@

\begin{figure}[t!]
\begin{subfigure}{0.6\textwidth}
<<sim_oneset1_mu, fig=TRUE, echo=FALSE>>=
plot_oneset(oneset1, only_dt = FALSE, only_df = FALSE, only_g = FALSE,
              compare_mu = TRUE, 
              add_dt = TRUE,
              add_df = TRUE,
              add_g = TRUE,
              nomodel = FALSE)
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.6\textwidth}
<<sim_oneset1_sigma, fig=TRUE, echo=FALSE>>=
  plot_oneset(oneset1, only_dt = FALSE, only_df = FALSE, only_g = FALSE,
              compare_mu = FALSE,
              compare_sigma_line = TRUE, 
              add_dt = TRUE,
              add_df = TRUE,
              add_g = TRUE,
              nomodel = TRUE)
@
\end{subfigure}
\caption{\label{plot_sim_oneset1} ... }
\end{figure}

\begin{figure}[t!]
\begin{subfigure}{0.6\textwidth}
<<sim_oneset10_mu, fig=TRUE, echo=FALSE>>=
plot_oneset(oneset10, only_dt = FALSE, only_df = FALSE, only_g = FALSE,
              compare_mu = TRUE, 
              add_dt = TRUE,
              add_df = TRUE,
              add_g = TRUE,
              nomodel = FALSE)
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.6\textwidth}
<<sim_oneset10_sigma, fig=TRUE, echo=FALSE>>=
  plot_oneset(oneset10, only_dt = FALSE, only_df = FALSE, only_g = FALSE,
              compare_mu = FALSE,
              compare_sigma_line = TRUE, 
              add_dt = TRUE,
              add_df = TRUE,
              add_g = TRUE,
              nomodel = TRUE)
@
\end{subfigure}
\caption{\label{plot_sim_oneset10} ... }
\end{figure}

In the case of smooth effects of the coefficients on the parameter function the GAMLSS is closest to the true parameter while the steps of the tree are cleary too rough. By combining various trees the forest can smooth out these steps, however, not as well as the GAMLSS.\\
Looking at the other extreme situation in which the parameter functions include abrupt changes the tree captures them very well wich also leads to a good fit of the forest model. In contrast to the two tree structured models the GAMLSS has its difficulties with these steps and therefore fluctuates around the true parameter function.\\
\\

Now, to get a more profound impression, the methods are evaluated on various data sets. The value of $\kappa$ is increasing from 1 to 10 by a stepsize of 1 and for each step ... data sets are generated.\\
\\
%\subsubsection{Means of Comparison}
In order to compare the performance of the chosen methods over various data sets, the continuous ranked probability score (CRPS) is calculated for each model on each data set.

%three different values are calculated for the resulting models: the root of the mean squared error (RMSE), the log-likelihood value and the continuous ranked probability score (CRPS). All of them are evaluated out of sample, e.g. on a part of the data set that has not been used to build the model.\\
%Therefore, overfitting would lead to worse results.\\
%\\
%The RMSE represents how close the estimated expected values are to the true observations. For a response variable $Y$ and a set of observations $\textbf{y} = \{y_i\}_{i = 1,\ldots,n}$ with the corresponding estimated parameters $\hat{\boldsymbol{\theta}} = \{\hat{\theta}_i\}_{i = 1,\ldots,n}$ the RMSE is defined as follows.

%$$
%\text{RMSE}(Y, \boldsymbol{y}, \boldsymbol{\hat{\theta}}) = \sqrt{\frac{1}{n} \sum_{i=1}^n (\mathbb{E}_{\hat{\theta}_i}(Y) - y_i)^2}
%$$

%In most cases this value does not measure the distributional fit as only the expected value is considered. However, in the case of a censored normal distribution the expected value of the observed variable is a transformation of the expected value of the latent variable which includes the location and the scale parameter.\\
%\\
%A typical way to evaluate the distributional fit is provided by the log-likelihood value. For each observation $y$ and the corresponding predicted parameter $\hat{\theta}$ the value of the log-likelihood function $\ell(\hat{\theta}; y)$ is calculated and then averaged over all observations in the testing data set. The higher the value the better the predictions of the model are.\\
%\\
%The CRPS is another value that measures the accuracy of a probabilistic model. For one observation $y$ with the predicted distribution function $F$ the CRPS value is calculated by
The CRPS measures the accuracy of a probabilistic model. For one observation $y$ with the predicted distribution function $F$ the CRPS value is calculated by
$$
\text{CRPS}(y, F) = \int_{-\infty}^\infty (F(z) - \mathbf{1}(y \leq z))^2 dz
$$
where \(\mathbf{1}\) is the indicator function. \\
In this simulation study the CRPS is evaluated out of sample, e.g. on a data set that has not been used to build the model but has been generated by the same process and with the same settings as the original data set. In the following this second data set on which the CRPS is evaluated is called testing data set while the original data set on which the models are built on is called learning data set.
The CRPS value is first calculated for each observation in the testing data set and then the average of these values is taken. As the CRPS can be seen as a generalization of the mean absolute error, a low value indicates a good fit. In applications, one big advantage in terms of interpretability is that the resulting value is on the scale of the observation. \\

%The methods are then compared by the out-of-sample RMSE for the location parameter $\mu$ and the shape parameter $\sigma$ as well as by %the out-of-sample log-likelihood of the fitted models, averaged within each step.
%(In that way it can be investigated how the results change for a changing value of $\kappa$.)

@@

%\begin{figure}[t!]
%\begin{subfigure}{0.6\textwidth}
%<<sim_rmse_exp, fig=TRUE, echo=FALSE>>=
%plot_rmse(simres, type = "exp")
%@
%\end{subfigure}
%\hspace{-1cm}
%\begin{subfigure}{0.6\textwidth}
%<<sim_rmse_par, fig=TRUE, echo=FALSE>>=
%plot_rmse(simres, type = "par")
%@
%\end{subfigure}
%\caption{\label{plot_sim_rmse}Results of the simulation study (RMSE).\\
%left: continuous lines represent the difference between the true and the estimated expected value, dotted lines represent the difference %between the observations and the estimated expected value\\
%right: continuous lines represent the location parameter, dotted lines the scale parameter}
%\end{figure}

\begin{figure}[t!]
\begin{center}
%\begin{subfigure}{0.6\textwidth}
<<sim_crps, fig=TRUE, echo=FALSE>>=
plot_crps(simres)
@
\caption{\label{plot_sim_crps}Results of the simulation study: CRPS} %\code{disttree} vs. \code{distforest} vs. \code{gamlss} vs. \code{randomForest} vs. %\code{bamlss} vs. \code{gamboostLSS}\\
%right: levelplot of the parameter function for $\kappa=5$}
\end{center}
\end{figure}

%\end{subfigure}
%\hspace{-1cm}
%\begin{subfigure}{0.6\textwidth}
%<<sim_levelplot, fig=TRUE, echo=FALSE>>=
%levelfun <- function(x) simres$fun(x,5)
%
%ncuts <- 100
%colR <- colorRampPalette(c("blue","green"))                                       
%cols <- colR(ncuts+1)
%griddata <- expand.grid(list(x1 = seq(-1, 1, 0.01), x4 = seq(0, 1, 0.2)))
%griddata$mu <- levelfun(griddata)
%levelplot(mu ~ x1 * x4, data = griddata, xlab = "x1", ylab = "x4", 
%          region = TRUE, cuts = ncuts, col.regions = cols)

%@
%\end{subfigure}
%\caption{\label{plot_sim_ll}left: results of the simulation study: log-likelihood\\ %\code{disttree} vs. \code{distforest} vs. \code{gamlss} vs. \code{randomForest} vs. %\code{bamlss} vs. \code{gamboostLSS}\\
%right: levelplot of the parameter function for $\kappa=5$}
%\end{figure}
